### 2016-07-01 sec6_experiment.tex
Final Version
Submitted to VLDB

\textbf{Datasets}: We use three real trajectory datasets in different application scenarios:
\begin{itemize}
\item{Shopping}\footnote{\url{http://www.irc.atr.jp/crest2010_HRI/ATC_dataset/}}. The dataset contains
  trajectories of visitors in the ATC shopping center in Osaka. To better capture the indoor activities, the visitor locations are sampled every half second, resulting in $13,183$ long trajectories. 
\item{GeoLife}~\footnote{\url{http://research.microsoft.com/en-us/projects/geolife/}}. The dataset essentially keeps all the travel records of 182 users for a period
of over three years, including multiple kinds of transportation modes (walking, driving and taking public
transportation). For each user, the GPS information is collected periodically and 91 percent of the trajectories
are sampled every 1 to 5 seconds.
\item{Taxi}. The dataset tracks the trajectories of $15,054$ taxies in Singapore over August 2008. The sampling
rate is around 30 seconds. 
\end{itemize}

\textbf{Preprocessing}: We replace timestamps with global sequences (starting from $1$) for each dataset. 
We set a fixed sampling rate for each dataset (i.e., Geolife = 5 seconds, Shopping=0.5 seconds, Taxi = 30 seconds)
and use linear interpolation to fill missing values.
%For each dataset, we set a fixed sampling rate (CLEARLY STATE THE SAMPLING RATE FOR EACH DATASET) and use linear interpolation to fill missing values. 
%Finally, we obtain a set of trajectories with equal length. 
%Finally, we obtain a set of dense trajectories with equal length. 
For the clustering method, we use DBSCAN~\cite{ester1996density} and customize its two parameters $\epsilon$ (proximity threshold) and $minPt$ (the minimum number of points required to form a dense region). We set $\epsilon=5$, $minPt=10$ for GeoLife and Shopping datasets; and $\epsilon=20$, $minPt=10$ for Taxi dataset. Note that other clustering methods or settings can also be applied. 
%The clustered snapshots are stored in HDFS as $\langle t, S_t \rangle$ pair, where $t$ is the timestamp, $S_t$ contains the clusters at snapshot $t$. 
After preprocessing, the statistics of the three datasets are listed in Table~\ref{exp:dataset}.

@@

	\begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_M.eps}
        \caption{Geolife vary $M$}
    \end{subfigure}
    \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_K.eps}
        \caption{Geolife vary $K$}
    \end{subfigure}
    \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_L.eps}
        \caption{Geolife vary $L$}
    \end{subfigure}
       \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_G.eps}
        \caption{Geolife vary $G$}
    \end{subfigure}       
 	 \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_o.eps}
        \caption{Geolife vary $O_r$}
    \end{subfigure}
 	\begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_t.eps}
        \caption{Geolife vary $T_r$}
    \end{subfigure}

@@

\caption{Performance of SPARE and TRPM on real datasets under different pattern parameters.}
\label{exp:performance_vary}
\end{figure*}
\textbf{Varying $M$}: Figures~\ref{exp:performance_vary} (a),(g),(m)
present the performance with increasing $M$. The SPARE framework demonstrates a clear superiority over the TRPM framework, with 
a performance gain by a factor of  $2.7$ times in Shopping, $3.1$ times in Geolife and
$7$ times in Taxi. As $M$ increases, the running time of both frameworks slightly improves because the number of clusters in each snapshot drops, generating fewer valid candidates.

\textbf{Varying $K$}: The performance with increasing $K$ is shown in Figure~\ref{exp:performance_vary} (b),(h),(n).  SPARE tends to run faster, whereas the performance of TRPM degrades dramatically. This is caused by the \emph{sequence simplification} procedure in SPARE, which can prune many candidates with large $K$. However, the line sweeping algorithm in TRPM does not utilize such property for pruning. It takes longer time because more replicated data has to be handled in each partition.

@@

\textbf{Varying $G$}: Figures~\ref{exp:performance_vary} (d),(j),(p) present the performances with increasing $G$.  TRPM is rather sensitive to $G$. When $G$ is relaxed to larger values, more valid patterns would be generated. TPRM has to set a higher replication factor and its running time degrades drastically when $G$ increases from $20$ to $30$. In contrast, with much more effective pruning strategy, our SPARE scales very well with $G$.

@@

\begin{table}[h]
\centering \small
\begin{tabular}{|l|c|c|c|c|}
\hline
\multirow{2}{*}{\textbf{Dataset}} & \multicolumn{2}{l|}{\textbf{SPARE-RD}} & \multicolumn{2}{l|}{\textbf{SPARE}} \\ \cline{2-5} 
                         & Straggler        & Std. Dev.     & Straggler       & Std. Dev.      \\ \hline
Shopping                 & 295              & 41         & 237             & 21       \\ \hline
GeoLife                  & 484              & 108        & 341             & 56       \\ \hline
Taxi                     & 681              & 147        & 580             & 96       \\ \hline
\end{tabular}
  \caption{Statistics of execution time among executors (in seconds).}
  \label{tbl:strags}
\end{table}

\subsubsection{Load balance}
To study the effect of load balance in the SPARE framework, we use random task allocation (the default setting of Spark) as a baseline, denoted by SPARE-RD, and compare it with our best-fit method. In best-fit, the largest unassigned star is allocated to the currently most lightly loaded reducer.
Figure~\ref{exp:wl}(a) shows the breakdown of the costs in the map-reduce stages for SPARE and SPARE-RD. We observe that the map and shuffle time of SPARE and SPARE-RD are identical. The difference is that SPARE incurs an additional overhead to generate an allocation plan for load balance (around $4\%$ of the total cost), resulting in significant savings in the reduce stage (around $20\%$ of the total cost). We also report the cost of \emph{straggler}, i.e., the longest job, and the standard deviation (Std. Dev.) for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.

\subsection{Scalability}
Finally, we examine the scalability of our proposed frameworks in two contexts. 
First, we use the single-node mode and examine the performance with multi-cores. 
We also compare it with the state-of-the-art solutions for swarm and platoon patterns 
since they were designed to run in a centralized server. 
Second, we examine the pattern detection time with increasing computing resources (number of machines). 
To support swarm and platoon, we sample 60\% of each dataset for this experiments. 
The results are reported in Figure~\ref{exp:scalability}. 
First, the centralized schemes are not suitable to discover patterns in 
large-scale trajectory databases. It takes nearly $30$ hours to 
detect \emph{swarm}s and $11$ hours to detect \emph{platoon}s in the Taxi dataset in a single machine. 
In contrast, when utilizing the multi-core environment, 
SPARE achieves $8$ times speed up compared to \emph{platoon} and $22$ times speed up compared to \emph{swarm}. 
%our TRPM and SPARE achieves 2.4 times and 7.1 times speedup respectively. 
Second, we see that both TRPM and SPARE demonstrate promising scalability in terms of the number of machines available. The running time almost decreases inversely as more machines are used. 
When there are $11$ nodes in the cluster, 
SPARE achieves at most $73$ times faster 
than \emph{platoon} and at most $170$ times better than \emph{swarm}.
%state-of-the-art \emph{platoon} detector in a centralized server.

\begin{figure*}[t]
\centering
\begin{subfigure}[b]{0.31\textwidth}
    \includegraphics[width=\textwidth]{/exp/spare/scalability-shopping.eps}
        \caption{Shopping vary $N$}
    \end{subfigure}
 	 \begin{subfigure}[b]{0.31\textwidth}
        \includegraphics[width=\textwidth]{/exp/spare/scalability-geolife.eps}
        \caption{Geolife vary $N$}
    \end{subfigure}
    	 \begin{subfigure}[b]{0.31\textwidth}
        \includegraphics[width=\textwidth]{/exp/spare/scalability-taxi.eps}
        \caption{Taxi vary $N$}
    \end{subfigure}
 \caption{Comparisons among TRMP, SPARE, PLATOON and SWARM.}
 \label{exp:scalability}
\end{figure*}

>>>>

\textbf{Datasets}: We use three real trajectory datasets in different application scenarios:
\begin{itemize}
\item{Shopping}\footnote{\url{http://www.irc.atr.jp/crest2010_HRI/ATC_dataset/}}: The dataset contains
  trajectories of visitors in the ATC shopping center in Osaka. To better capture the indoor activities, the visitor locations are sampled every half second, resulting in $13,183$ long trajectories. 
\item{GeoLife}~\footnote{\url{http://research.microsoft.com/en-us/projects/geolife/}}: The dataset essentially keeps all the travel records of 182 users for a period
of over three years, including multiple kinds of transportation modes (walking, driving and taking public
transportation). For each user, the GPS information is collected periodically and 91 percent of the trajectories
are sampled every 1 to 5 seconds.
\item{Taxi}~\footnote{Taxi is our proprietary dataset}: The dataset tracks the trajectories of $15,054$ taxies in Singapore. For each taxi, the GPS information are continually collected for one entire month with the sampling rate around 30 seconds.
\end{itemize}

\textbf{Preprocessing}: We replace timestamps with global sequences (starting from $1$) for each dataset. 
We set a fixed sampling rate for each dataset (i.e., GeoLife = 5 seconds, Shopping=0.5 seconds, Taxi = 30 seconds)
and use linear interpolation to fill missing values.
%For each dataset, we set a fixed sampling rate (CLEARLY STATE THE SAMPLING RATE FOR EACH DATASET) and use linear interpolation to fill missing values. 
%Finally, we obtain a set of trajectories with equal length. 
%Finally, we obtain a set of dense trajectories with equal length. 
For the clustering method, we use DBSCAN~\cite{ester1996density} and customize its two parameters $\epsilon$ (proximity threshold) and $minPt$ (the minimum number of points required to form a dense region). We set $\epsilon=5$, $minPt=10$ for GeoLife and Shopping datasets; and $\epsilon=20$, $minPt=10$ for Taxi dataset. Note that other clustering methods or settings can also be applied. 
%The clustered snapshots are stored in HDFS as $\langle t, S_t \rangle$ pair, where $t$ is the timestamp, $S_t$ contains the clusters at snapshot $t$. 
After preprocessing, the statistics of the three datasets are listed in Table~\ref{exp:dataset}.

@@

	\begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_M.eps}
        \caption{GeoLife vary $M$}
    \end{subfigure}
    \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_K.eps}
        \caption{GeoLife vary $K$}
    \end{subfigure}
    \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_L.eps}
        \caption{GeoLife vary $L$}
    \end{subfigure}
       \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_G.eps}
        \caption{GeoLife vary $G$}
    \end{subfigure}       
 	 \begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_o.eps}
        \caption{GeoLife vary $O_r$}
    \end{subfigure}
 	\begin{subfigure}[b]{0.16\textwidth}
        \includegraphics[width=\textwidth]{/exp/performance/geolife_vary_t.eps}
        \caption{GeoLife vary $T_r$}
    \end{subfigure}

@@

\caption{Performance of SPARE and TRPM on real datasets under different pattern parameters.}
\label{exp:performance_vary}
\end{figure*}
\textbf{Varying $M$}: Figures~\ref{exp:performance_vary} (a),(g),(m)
present the performance with increasing $M$. The SPARE framework demonstrates a clear superiority over the TRPM framework, with 
a performance gain by a factor of  $2.7$ times in Shopping, $3.1$ times in GeoLife and
$7$ times in Taxi. As $M$ increases, the running time of both frameworks slightly improve because the number of clusters in each snapshot drops, generating fewer valid candidates.

\textbf{Varying $K$}: The performance with increasing $K$ is shown in Figure~\ref{exp:performance_vary} (b),(h),(n).  SPARE tends to run faster, whereas the performance of TRPM degrades dramatically. This is caused by the \emph{sequence simplification} procedure in SPARE, which can prune many candidates with large $K$. However, the line sweep algorithm in TRPM does not utilize such property for pruning. It takes longer time because more replicated data has to be handled in each partition.

@@

\textbf{Varying $G$}: Figures~\ref{exp:performance_vary} (d),(j),(p) present the performances with increasing $G$.  TRPM is rather sensitive to $G$. When $G$ is relaxed to larger values, more valid patterns would be generated. TPRM has to set a higher replication factor and its running time degrades drastically when $G$ increases from $20$ to $30$. In contrast, with much more effective pruning strategy, our SPARE scales well with $G$. Particularly, SPARE is 14 times faster than TRPM when $G=20$ in GeoLife dataset.

@@

\begin{table}[h]
\centering \small
\begin{tabular}{|l|c|c|c|c|}
\hline
\multirow{2}{*}{\textbf{Dataset}} & \multicolumn{2}{l|}{\textbf{SPARE-RD}} & \multicolumn{2}{l|}{\textbf{SPARE}} \\ \cline{2-5} 
                         & Straggler        & Std. Dev.     & Straggler       & Std. Dev.      \\ \hline
Shopping                 & 295              & 41         & 237             & 21       \\ \hline
GeoLife                  & 484              & 108        & 341             & 56       \\ \hline
Taxi                     & 681              & 147        & 580             & 96       \\ \hline
\end{tabular}
  \caption{Statistics of execution time (in seconds) among executors.}
  \label{tbl:strags}
\end{table}

\subsubsection{Load balance}
To study the effect of load balance in the SPARE framework, we use random task allocation (the default setting of Spark) as a baseline, denoted by SPARE-RD, and compare it with our best-fit method. In best-fit, the largest unassigned star is allocated to the currently most lightly loaded reducer.
Figure~\ref{exp:wl} (a) shows the breakdown of the costs in the map-reduce stages for SPARE and SPARE-RD. We observe that the map and shuffle time of SPARE and SPARE-RD are identical. The difference is that SPARE incurs an additional overhead to generate an allocation plan for load balance (around $4\%$ of the total cost), resulting in significant savings in the reduce stage (around $20\%$ of the total cost). We also report the cost of \emph{straggler}, i.e., the longest job, and the standard deviation (Std. Dev.) for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.

\subsection{Scalability}
Finally, we examine the scalability of our proposed frameworks in two contexts. 
First, we use the single-node mode and examine the performance with multi-cores. 
We also compare it with the state-of-the-art solutions for swarm and platoon patterns 
since they were designed to run in a centralized server. 
Second, we examine the pattern detection time with increasing computing resources (number of machines). 
As the poor performances of swarm and platoon, we only use 60\% of each dataset for experiments. 
The results are reported in Figure~\ref{exp:scalability}. 
First, the centralized schemes are not suitable to discover patterns in 
large-scale trajectory databases. It takes nearly $30$ hours to 
detect \emph{swarm}s and $11$ hours to detect \emph{platoon}s in the Taxi dataset in a single machine. 
In contrast, when utilizing the multi-core environment, 
SPARE achieves $9$ times speed up compared to \emph{platoon} and $22$ times speed up compared to \emph{swarm}. 
%our TRPM and SPARE achieves 2.4 times and 7.1 times speedup respectively. 
Second, we see that both TRPM and SPARE demonstrate promising scalability in terms of the number of machines available. The running time almost decreases inversely as more machines are used. 
When there are $11$ nodes in the cluster, 
SPARE achieves at most $63$ times faster 
than \emph{platoon} and at most $170$ times better than \emph{swarm}.
%state-of-the-art \emph{platoon} detector in a centralized server.

\begin{figure*}[t]
\centering
\begin{subfigure}[b]{0.31\textwidth}
    \includegraphics[width=\textwidth]{/exp/spare/scalability-shopping.eps}
        \caption{Shopping vary $N$}
    \end{subfigure}
 	 \begin{subfigure}[b]{0.31\textwidth}
        \includegraphics[width=\textwidth]{/exp/spare/scalability-geolife.eps}
        \caption{GeoLife vary $N$}
    \end{subfigure}
    	 \begin{subfigure}[b]{0.31\textwidth}
        \includegraphics[width=\textwidth]{/exp/spare/scalability-taxi.eps}
        \caption{Taxi vary $N$}
    \end{subfigure}
 \caption{Comparisons among TRMP, SPARE, PLATOON and SWARM.}
 \label{exp:scalability}
\end{figure*}

### 2016-07-01 sec7_conclusion.tex
Final Version
Submitted to VLDB

\section{Conclusion and Future Work}
\label{sec:concl}

>>>>

\section{Conclusions and Future Work}
\label{sec:concl}
In this paper, we proposed a generalized co-movement pattern to unify those proposed in the past literature. We also devised two types of parallel frameworks in Spark that can scale to support pattern detection in trajectory databases with hundreds of millions of points. The efficiency and scalability were verified by extensive experiments on three real datasets. In the future, we intend to examine co-movement pattern detection in streaming data for real-time monitoring. How to extend the current parallel frameworks to support other types of advanced patterns is also of our interest.
